\subsection{Performance Management for Big Data Applications in Mobile Cloud Computing}

In mobile computing environment, the computation intensive and data intensive applications usually lead to unsatisfactory performance on resource limited mobile devices. As a new paradigm, cloud computing can enhance computing and storage capabilities of mobile devices for big data applications. The basic idea of computation offloading is to move the task execution from mobile devices to cloud infrastructures, which have clusters with faster processing speed.

Mobile computation and data offloading has been regarded as a research focus by researchers in recent years. The existing work on this problem differs in their cost-benefit models. The following factors are considered, including energy consumption of mobile devices, data transmission delay, application response time, etc. Although offloading optimization has been studied recently, some critical issues on this problem remain unsolved. Actually, only a few studies take unstable wireless bandwidth into consideration, which can be regarded as a primary constraint for the performance of offloading process in real environment.

In this research, we are to build the cost-benefit models and manage the performance of mobile offloading processes. Our objective is to minimize the cost and maximize the benefit. We will provide offloading strategies for multiple mobile users and help them achieve high performance. Moreover, we are to make offloading decisions dynamically by estimating available bandwidth of each mobile user. As a step forward, we will try to manage the bandwidth to improve the performance of big data applications in mobile cloud computing.

For this topic, we mainly have the following research objectives:

\begin{itemize}
  \item Provide offloading strategies for multiple mobile users and help them achieve high performance
  \item Design algorithms for dynamical offloading by estimating available bandwidth of each mobile user
  \item Propose methods for improving the performance of big data applications in mobile cloud computing
\end{itemize}
